Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms

### 2017-11-26 paper/paper.tex
Eight threaded freaks
Ruth's comments

A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
graph which \emph{preserves adjacency}---that is, if it maps adjacent vertices to adjacent vertices.
The isomorphism is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
non-adjacency. When working with labelled graphs, a subgraph isomorphism must also preserve labels, and
on directed graphs, it must preserve orientation. A \emph{common induced subgraph} of two graphs $G$
and $H$ is a pair of induced subgraph isomorphisms from a pattern graph $P$, one to $G$ and one to
$H$. A \emph{maximum common induced subgraph} is one with as many vertices as possible. (The
\emph{maximum common partial subgraph} problem is non-induced, with as many edges as
possible; this paper discusses only induced problems.)
A common variant of the problem requires a largest \emph{connected} subgraph
\cite{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.

@@

\subsection{Domain Splitting (McSplit)}

@@

@@

This algorithm is aimed primarily at large instances, where the two graphs are of different orders,
and where it is expected that the solution will involve most of the smaller graph (that is, $k$ is
expected to be low). The sequential implementation we work with does not support labels or the
connected variant.

@@

Most of the benchmark instances we will use come from a standard database for maximum common
subgraph problems \cite{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}. This benchmark
set can be used in a number of ways, for different variants of the problem. Following other recent
work \cite{DBLP:conf/cp/McCreeshNPS16,DBLP:conf/aaai/HoffmannMR17,o:McCreeshPT17}, we use it to
create five families of instance, as follows:

@@

With one exception, each plot gives the same conclusion: if we are working with a timeout of at
least $10^2$ milliseconds, then for any problem family and any sequential algorithm, if given the
option of switching to the corresponding parallel algorithm, then we should do so. For the McSplit
algorithm on both labelled, connected instances, the parallel algorithm does not quite catch up to
the sequential algorithm.

@@

A more principled approach is given in the second column of \cref{figure:cliquespeedups}. For
instances where the sequential run both succeeded and took at least 500 milliseconds, we plot the
distribution of speedups obtained. These histograms confirm our informal observations. However,
these plots are still not especially satisfactory: in order to calculate a speedup, we can only
consider instances where the sequential algorithm succeeded, and so these plots therefore underestimate
superlinear speedups. The choice of a 500 millisecond minimum sequential runtime is also rather
arbitrary, and is acceptable only if we expect the parallel algorithms will only be used on
relatively hard instances.

In the third column we show performance profiles \cite{DBLP:journals/mp/DolanM02}. A performance
profile is a cumulative plot of the relative performance of the parallel algorithm to the sequential
algorithm---instances occurring to the left of the vertical $x = 1$ line are speedups. Each plot
shows three options as different lines. The `all' line include easy instances whose sequential
runtime is below 500 milliseconds, whilst the other two lines include them. The `hard' line treats
sequential timeouts as having been solved at the time limit, whilst the `par10' line treats timeouts
as taking ten times longer than the timeout (this convention is common in portfolios).

Unfortunately, these three lines can paint very different pictures. For example, for unlabelled
instances on the top row, if we include easy instances, it appears that parallelism is better on
just over half of the instances, whilst if we exclude them, parallelism is always beneficial. If we
do not use the par10 scheme, the performance profile also suggests that there are around 500
instances where the speedup is below 10, whilst using par10 correctly shows that such instances are
rare. However, par10 is only effective in this regard because the ``typical'' speedup is in the
region of 10 (and this is a particular inconvenience because we seek a way of characterising
speedups which does not rely upon us already knowing that 10 is a reasonable choice of penalty).

@@

    \caption{Per-instance speedups from the clique algorithm on vertex- and edge-labelled, directed
    instances, when going from sequential to two threads in the first plot, then increasing the
    number of threads in subsequent plots. The final plot shows 32 threads versus a repeated run
    also with 32 threads. ?? 8 threaded results still running}\label{figure:cliquescale}
\end{figure}

>>>>

A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
graph which \emph{preserves adjacency}---that is, if it maps adjacent vertices to adjacent vertices.
The isomorphism is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent
vertices, preserving non-adjacency as well. When working with labelled graphs, a subgraph
isomorphism must preserve labels, and on directed graphs, it must preserve orientation. A
\emph{common induced subgraph} of two graphs $G$ and $H$ is a pair of induced subgraph isomorphisms
from a pattern graph $P$, one to $G$ and one to $H$. A \emph{maximum common induced subgraph} is one
with as many vertices as possible. (The \emph{maximum common partial subgraph} problem is
non-induced, with as many edges as possible; this paper discusses only induced problems.) A common
variant of the problem requires a largest \emph{connected} subgraph
\cite{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.

@@

\subsection{Domain Splitting (McSplit and McSplit$\downarrow$)}

@@

The McSplit$\downarrow$ algorithm is a variant designed for instances where we expect nearly all of
the smaller graph to be found. It branches first on result size, from largest possible result
downwards.

@@

This algorithm is aimed primarily at large instances, where the two graphs are of different orders,
and where it is expected that the solution will involve most of the smaller graph (that is, $k$ is
expected to be low). The sequential implementation we start with does not support labels or the
connected variant.

@@

Most of the benchmark instances we will use come from a standard database for maximum common
subgraph problems \cite{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}. This benchmark
set can be used in a number of ways, for different variants of the problem. Following other recent
work \cite{DBLP:conf/cp/McCreeshNPS16,DBLP:conf/aaai/HoffmannMR17,o:McCreeshPT17}, we use it to
create five families of instances, as follows:

@@

With one exception, each plot gives the same conclusion: if we are working with a solving time of at
least $10^2$ milliseconds, then for any problem family and any sequential algorithm, if given the
option of switching to the corresponding parallel algorithm, then we should do so. For the McSplit
algorithm on both labelled, connected instances, the parallel algorithm does not quite catch up to
the sequential algorithm.

@@

A more principled approach is given in the second column of \cref{figure:cliquespeedups}. For
instances where the sequential run both succeeded and took at least 500 milliseconds, we plot the
distribution of speedups obtained. These histograms confirm our informal observations. However,
these plots are still not especially satisfactory: in order to calculate a speedup, we can only
consider instances where the sequential algorithm succeeded, and so these plots underestimate
superlinear speedups. The choice of a 500 millisecond minimum sequential runtime is also rather
arbitrary, and is acceptable only if we expect the parallel algorithms will only be used on
relatively hard instances.

In the third column we show performance profiles \cite{DBLP:journals/mp/DolanM02}. A performance
profile is a cumulative plot of the relative performance of the parallel algorithm to the sequential
algorithm---instances occurring to the left of the vertical $x = 1$ line are speedups. Each plot
shows three options as different lines. The `all' line include easy instances whose sequential
runtime is below 500 milliseconds, whilst the other two lines include them. The `hard' line treats
sequential timeouts as having been solved at the time limit, whilst the `PAR10' line treats timeouts
as taking ten times longer than the timeout (this convention is common in portfolios
\cite{DBLP:conf/aaai/XuHL10}).

Unfortunately, these three lines can paint very different pictures. For example, for unlabelled
instances on the top row, if we include easy instances, it appears that parallelism is better on
just over half of the instances, whilst if we exclude them, parallelism is always beneficial. If we
do not use the PAR10 scheme, the performance profile also suggests that there are around 500
instances where the speedup is below 10, whilst using PAR10 correctly shows that such instances are
rare. However, PAR10 is only effective in this regard because the ``typical'' speedup is in the
region of 10 (and this is a particular inconvenience because we seek a way of characterising
speedups which does not rely upon us already knowing that 10 is a reasonable choice of penalty).

@@

    \caption{Per-instance speedups from the clique algorithm on vertex- and edge-labelled, directed
    instances, when going from sequential to two threads in the first plot, then increasing the
    number of threads in subsequent plots. The final plot shows 32 threads versus a repeated run
    also with 32 threads.}\label{figure:cliquescale}
\end{figure}
